---
title: "606 Assignment 3"
author: "Brett Burk"
date: "Wednesday, February 25, 2015"
output: pdf_document
---
\textbf{4.24.i} Using the definition of variance:
$$V[X] = E[(X- \mu)^2] = ((1-3.5)^2 + (2-3.5)^2 + (3-3.5)^2 + (4-3.5)^2 + (5-3.5)^2 + (6-3.5)^2)/6 = (6.25+2.25+.25)/3 = 2.91 \bar{6}$$
\textbf{4.24.ii} Using $V[X] = E[X^2] - E[X]^2$:
$$((1+4+9+16+25+36)/6)-3.5^2 = 2.91 \bar{6}$$ I prefer the second because it is computationally simpler. \newline
\textbf{4.36}
$$Var[X+Y] = Var[X] + Var[Y] + 2 Cov(X,Y)$$
$$Cov(X,Y) = -0.5\sigma_x \sigma_y$$
$$Var[X+Y] = var[X] + Var[Y] - 1 \sigma_x \sigma_y$$
Assuming by 'the same distribution' it is meant that mean(x) = mean(y) and var(x) = var(y):
$$Var[X+y]=2var[X] - var[x] = Var[X] = Var[Y]$$

\textbf{4.40.a} $$V[3X^2 - Y] = 9V[X^2] + V[Y]$$
$$9(E[X^4] - E[X^2]^2) + (E[Y^2] - E[Y]^2)$$
$$9(15-4)+(6-4)=101$$
\textbf{4.40.b} $$Cov(X,Y) = E[XY] - E[X]E[Y]$$
As they are independent
$$0 = E[XY] - E[X]E[Y]$$
$$E[X]E[Y] = E[XY]$$
$$15*14 = 1410$$
\textbf{4.40.c} $$Cov(X,X^2) = E[X^3] - E[X]E[X^2]$$
$$5-1*2 = 3$$
\textbf{4.40.d} $$V[X^2 Y^2] = E[X^4 Y^4] - (E[X^2]E[Y^2])^2=E[X^4]E[Y^4]-(E[X^2]E[Y^2])^2=15 \times 94 - (2 \times 6)^2 = 1266$$
$$15 \times 94 - 4 \times 6 = 1386$$
\textbf{4.56}
```{r}
fourthree <- function(){
  roll <- sample(1:6,1)
  if(roll <= 3){
    return(-10)
  }
  if(roll %in% c(4,5)){
    return(0)
  }
  else{
    return(14)
  }
}
ft <- vector(length = 10000)
ft <- sapply(1:10000, function(x) fourthree())
expectation <- sum(ft)/length(ft)
expectation
variance <- sum(ft^2)/length(ft) - expectation^2
variance
```
(I shifted the answer on this one, as my original assumed 'winning $24' meant getting back your money in addition to the $24)
\textbf{4.58}
```{r}
x.dist <- rpois(100000,1)
y.dist <- rpois(100000,2)
```
4.40.a:
```{r}
var(3*(x.dist^2) - y.dist)
```
4.40.b:
```{r}
mean(x.dist^4 * y.dist^4)
```
4.40.c:
```{r}
cov(x.dist, x.dist^2)
```
4.40.d:
```{r}
var(x.dist^2 * y.dist^2)
```
\textbf{5.4.a}
```{r}
 0.99^109 * .01^1
```
\textbf{5.4.b}
```{r}
.99^109
```
\textbf{5.4.c} \newline
100
\newline

\textbf{5.12}
$$
E[K] = \frac{k}{4}({k \choose 4} p^4 (1-p)^{k - 4} + {k \choose 4} (1 - p)^4 p^{k - 4})  \text{ for } k = 4,5,6,7
$$
\textbf{5.20.a} $$\mu_s = \frac{\text{tagged} \times \text{caught}}{\text{total}}=2$$
$$\sigma_s = \frac{n(N-n)r(N-r)}{N^2(N-1)} \approx 1.73146$$
\textbf{5.20.b}
```{r}
little.n <- 20

big.n <- 500
big.k <- 50

temp.p <- 1
for(k in 0:2){
  temp.p <- temp.p - ((choose(big.k, k) * choose(big.n - big.k, little.n - k))/choose(big.n, little.n))
  
}
temp.p
```
\textbf{5.28.a}
$$Cov(X,Y)_n = n(P_{Both} - P_{X}P_{Y}) = 10(0-\frac{1}{12}) = \frac{-5}{5}$$
\textbf{5.28.b} $$10(\frac{1}{6} - \frac{1}{12}) = \frac{5}{6}$$
\textbf{5.32} Assuming that homework could be given on the last class, we have that homework could be given on classes 2,3,5,7,11,13,17,19,23,29, thus we have $\frac{14}{42}$. \newline
Let X be the number of days with no homework, with $n=42$ and $p=\frac{1}{3}$ then it has a $\mu$ of 14. \newline \newline
\textbf{5.34} This is a hypergeometric distribution with $N=30, n=10, K=6, and k=3$ thus we have $$\frac{{6 \choose 3}{24 \choose 7}}{{30 \choose 10}} = 0.23039$$
\textbf{5.38}
```{r}
set.seed(01191984)
coupon <- function(n,p){
  have <- array(FALSE, dim=n)
  count <- 0
  while (sum(have) < n)
    {
      picked <- sample(1:n, 1, prob=p, replace=TRUE)
      have[picked] <- TRUE
      count <- count + 1
    }
  return(count)
}

p <- dbinom(0:6, 6, 0.5)
# p <- dbinom(0:52, 52, 0.5)
coupon(length(p), p)

curr.counts <- array(0, dim=10000)
curr.counts <- sapply(1:length(curr.counts), function(x) coupon(7,p))
mean(curr.counts)
sd(curr.counts)



```
\textbf{5.39}
```{r}
# Returns TRUE if they won the series
series.calc <- function(prob) {
  series.samp <- rbinom(4, 1, prob)
  while(length(series.samp) < 7 && max(table(series.samp)) < 4){
    series.samp <- c(series.samp, rbinom(1,1,prob))
  }
  return(sum(series.samp)/length(series.samp) > 0.5)
}

prob <- 0.25
series.low.counts <- array(0, dim=10000)
series.low.counts <- sapply(1:length(series.low.counts), function(x) series.calc(prob))
mean(series.low.counts)
sd(series.low.counts)

prob <- 0.6
series.high.counts <- array(0, dim=10000)
series.high.counts <- sapply(1:length(series.high.counts), function(x) series.calc(prob))
mean(series.high.counts)
sd(series.high.counts)
