---
title: "Collaboration1"
author: "Brett Burk"
date: "Sunday, February 22, 2015"
output: pdf_document
---
Just change the wd to the appropriate directory for your machine
```{r}
wd <- "C:\\Users\\Brett\\Dropbox\\CUNY\\606\\Week2\\"
```
Loading in:
```{r}
sales <- read.csv(paste0(wd, 'sales.csv'))
head(sales)
details <- read.csv(paste0(wd, 'details.csv'))
details
summary(sales)
dem <- sales[c('demand.veggie', 'demand.ham', 'demand.turkey')]
```
\textbf{Veggie:}
```{r, fig.height = 3, fig.width=6}
v <- dem$demand.veggie
print('veggie')
IQR(v)
sd(v)
hist(v)
hist(rpois(130, mean(v)))
boxplot(v)
```
\textbf{Turkey:}
```{r, fig.height = 3, fig.width=6}
t <- dem$demand.turkey
IQR(t)
sd(t)
hist(t)
hist(rpois(130, mean(t)))
boxplot(t)
```
\textbf{Ham:}
```{r, fig.height = 3, fig.width=6}
h <- dem$demand.ham
IQR(h)
sd(h)
hist(h)
hist(rpois(130, mean(h)))
boxplot(h)
```
\textbf{All three combined}
```{r}
vth <- v+t+h
hist(rpois(130, vth))
summary(vth)
boxplot(vth)
rpois(130, vth)
```
new work: \newline
\textbf{Sales ratios:}
```{r}
sales['vt.ratio'] <- sales$demand.veggie/sales$demand.turkey
mean(sales$vt.ratio)
sd(sales$vt.ratio)
sales['vh.ratio'] <- sales$demand.veggie/sales$demand.ham
mean(sales$vh.ratio)
sd(sales$vh.ratio)
sales['th.ratio'] <- sales$demand.turkey/sales$demand.ham
mean(sales$th.ratio)
sd(sales$th.ratio)
```
Well that seems like a dead end, the ratios vary wildly, 
```{r}
> sum(available.ham) * 2.50
Error: object 'available.ham' not found
> sum(sales$available.ham) * 2.50
[1] 5125
> sum(sales$available.turkey) * 2.50
[1] 5600
> sum(sales$available.ham) * 3
[1] 6150
> sum(sales$available.veggie) * 2.5
[1] 3075
> 5600 + 6150 + 3075
[1] 14825
> (5600 + 6150 + 3075)/130
[1] 114.0385
> (5600 + 6150 + 3075)/131
[1] 113.1679


demand.df <- cbind(sales$demand.ham - 15, sales$demand.turkey - 20, sales$demand.veggie - 13)
cost.df <- c(15 * 130, 20 * 130, veggie * 130
      )